Applied Interpretable Inference: Connecting Probabilistic Programming with Epidemiology Simulators

### 2019-04-23 paper/main.tex
updated

Simulators encode several years, if not decades, of research and development and so by their very nature are structurally complicated. Since the development happens over several decades, usually without documentation, understanding a code base written with legacy libraries and software is incredibly difficult and an almost impossible task for a non-expert, or expert who has not had access to the previous years of development. Our method of hijacking simulators builds on the work of~\cite{baydin2018efficient} and extends the framework to encapsulate a more diverse range of simulators, in particular epidemiology simulators. Our framework provides a simple solution and makes it easy to transform arbitrary stochastic simulators into probabilistic programs, regardless of the complexity of the simulator and the language that the simulator is written in\footnote{The framework currently supports 9 popular languages, but there is nothing to stop this being extended to any language of interest.}. In addition to this, we can understand the structure of programs in simulators for both decoding \emph{black-box structures} and for posterior inference. This enables two things. 1) It enables software developers and researchers to understand complicated code bases. 2) It provides interpretable inference results that provide policy makers with predictions that are truly interpretable, in the sense that the end-user of the inference results understands what physical events led to inference outcome. This is critical in several domains, and is particularly important in the medical domain.

@@

In order to hijack the stochastic simulator we are only required to override the location of the stochastic primitives, i.e. the operations that generate the stochasticity and randomness within the simulator. We will walk through how this is done for EMOD and OpenMalaria,  you will see that it is procedurally identical, however, given the sheer number of simulators it maybe slightly different procedurally. To our knowledge it shouldn't be, but we have not been able view of all simulators.

>>>>

Simulators encode several years, if not decades, of research and development and so by their very nature are structurally complicated. Since the development happens over several decades, usually without documentation, understanding a code base written with legacy libraries and software is incredibly difficult and an almost impossible task for a non-expert, or expert who has not had access to the previous years of development. Our method of hijacking simulators builds on the work of~\cite{baydin2018efficient} and extends the framework to encapsulate a more diverse range of simulators, in particular epidemiology simulators. Our framework provides a simple solution and makes it easy to transform arbitrary stochastic simulators into probabilistic programs, regardless of the complexity of the simulator and the language that the simulator is written in\footnote{The framework currently supports 9 popular languages, but there is nothing to stop this being extended to any language of interest.}. In addition to this, we can understand the structure of programs in simulators for both decoding \emph{black-box structures} and for posterior inference. This enables two things. 1) It enables software developers and researchers to understand complicated code bases. 2) It provides interpretable inference results that provide policy makers with predictions that are truly interpretable, in the sense that the end-user of the inference results understands what physical events led to inference outcome. This is critical in several domains, and is particularly important in the medical domain. We provide examples of both in Section~\textbf{TO ADD REF}.

@@

In order to hijack the stochastic simulator we are only required to override the location of the stochastic primitives, i.e. the operations that generate the stochasticity and randomness within the simulator. We will walk through how this is done for EMOD and OpenMalaria,  you will see that it is procedurally identical, however, given the sheer number of simulators it maybe slightly different procedurally. To our knowledge it shouldn't be, but we have not been able view of all stochastic simulators. 
The first step is to build a containerized environment, such as Docker\textbf{ADD citation} and Singularity, which enables the simulators to be
run on any device, independent of different hardware architectures removing the requirement for hardware specific machines. We provide examples of building a containerized environment in the Supplement \textbf{ADD ref}. This environment only has to be built once, which then enables the simulator to be run simultaneously on multiple machines. 
The second step
